Slight Positive Correlation Between Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-25 outputs/paper/paper.qmd
update graphs and bubble charts

This paper thus inspects gender and racial identities of all Toronto police arrests from 2020 to 2021, and assesses if there is a correlation between marginalized identities and police strip searching. Foremost, this paper aims to be a form of data activism to counter the injustices related to police power abuse around strip searching of the vulnerable Toronto population. To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using bar graphs and various other graphs. Particularly, for each category of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching. With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd]. Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021. Yet, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].

@@

#referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
head(cleaned_race_gender_data, 10) |>
  kable(
    col.names = c("ID", "Race", "Gender", "Strip Search", "Items Found"),
    booktabs = TRUE
  )
```

@@

Except for "ID" column, all other columns in this new dataset are reasons which account for why the Toronto police force conducts strip search under certain arrests. For instance, "Reason - Assist Escape" is for people who may have helped other suspects escape and "Reason - Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi. For each "Reason" column, if the data cell contains `0`, this means that strip search doesn't happen because of that particular reason. Vice versa, if it contains `1`, this means strip search happens because of that reason. In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()`.

@@

  head(cleaned_search_reasons_data, 10) |>
  select(-c(race, gender)) |>
  kable(
    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items_Found", "Total Reasons"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options="scale_down")
```

@@

cleaned_search_reasons_data |>
  group_by(total_reasons, items_found) |>
  count() |>
  rename(count = n) |>
  kable(
    col.names = c("Number of Reasons", "Items_Found", "Count"),
    booktabs = TRUE
  )
```

```{r}
#| echo: false
#| message: false
#| label: fig-scatter-plot-reasons
#| fig-cap: Relationship Between Reasons for and Items Found During Per Strip Search in Toronto, 2020-2021

#create a scatterplot to show each observation in the dataset, comparing the number of reasons to the items found in each strip search
#colour the points based on the police division where the group of shootings occurred
#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
cleaned_search_reasons_data |>
  group_by(total_reasons, items_found, race) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = race)) +
  geom_point() +
  theme_bw() +
  theme(legend.key.size = unit(0.5, 'cm'))+
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Race"
       ) +
  scale_fill_viridis_d(option="cividis")

cleaned_search_reasons_data |>
  group_by(total_reasons, items_found, race) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = race)) +
  geom_point() +
  theme_bw() +
  theme(legend.key.size = unit(0.5, 'cm'))+
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Race"
       ) +
  scale_fill_viridis_d(option="cividis") 
```

## Discussion

>>>>

This paper thus inspects gender and racial identities of all Toronto police arrests from 2020 to 2021, and assesses if there is a correlation between marginalized identities and police strip searching. Foremost, this paper aims to be a form of data activism to counter the injustices related to police power abuse around strip searching of the vulnerable Toronto population. To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked bar charts and bubble charts. Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching for intersecting social identities With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd]. Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021. Yet, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].

@@

#referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
head(cleaned_race_gender_data, 6) |>
  kable(
    col.names = c("ID", "Race", "Gender", "Strip Search", "Items Found"),
    booktabs = TRUE
  )
```

@@

Except for "ID" column, all other columns in this new dataset are reasons which account for why the Toronto police force conducts strip search under certain arrests. For instance, "Reason - Assist Escape" is for people who may have helped other suspects escape and "Reason - Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi. For each "Reason" column, if the data cell contains `0`, this means that strip search doesn't happen because of that particular reason. Vice versa, if it contains `1`, this means strip search happens because of that reason. In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.

@@

  head(cleaned_search_reasons_data, 6) |>
  select(-c(race, gender)) |>
  kable(
    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items_Found", "Total Reasons"),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options="scale_down")
```

@@

cleaned_search_reasons_data |>
  group_by(total_reasons, items_found) |>
  count() |>
  rename(count = n) |>
  kable(
    col.names = c("Number of Reasons", "Items Found", "Count"),
    booktabs = TRUE
  )
```

```{r}
#| echo: false
#| message: false
#| label: fig-reasons-race
#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021

#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
#colour the points based on the perceived race of the captive
#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
cleaned_search_reasons_data |>
  group_by(total_reasons, items_found, race) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = race, size = count)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  theme(legend.key.size = unit(0.5, 'cm'))
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Race"
       ) +
  scale_fill_viridis_d(option="cividis")

```{r}
#| echo: false
#| message: false
#| label: fig-reasons-gender
#| fig-cap: Relationship Between Reasons for Strip Search and Gender in Toronto, 2020-2021

cleaned_search_reasons_data |>
  group_by(total_reasons, items_found, gender) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = gender, size = count)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  theme(legend.key.size = unit(0.5, 'cm'))
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Gender"
       ) +
  scale_fill_viridis_d(option="cividis")

```
## Discussion
